### 2020-09-02 R/spectral-lib.Rmd
adding Max's gist code for extracting the spectral data

---
title: "Extracting all spectra for NEON mapped stems"
author: "Max Joseph"
date: "2020-06-11"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, message=FALSE}
library(tidyverse)
library(neonUtilities)
#library(devtools)
#install_github('NEONScience/NEON-geolocation/geoNEON', dependencies=TRUE)
library(geoNEON)
library(sp)
#devtools::install_github('earthlab/neonhs')
library(neonhs)
library(Rtsne)
library(ggthemes)
```

Acquire vegetation data:

```{r get-veg-data}
dir.create('data')
zipsByProduct(dpID = "DP1.10098.001",
              savepath = "data",
              check.size = FALSE)
stackByTable("data/filesToStack10098", folder = TRUE)
```

Read the mapping data and apparent individual data:

```{r read-map}
vegmap <- read_csv("data/filesToStack10098/stackedFiles/vst_mappingandtagging.csv") %>%
  filter(!is.na(stemAzimuth), !is.na(pointID))
vegind <- read_csv("data/filesToStack10098/stackedFiles/vst_apparentindividual.csv")
```

Get precise locations for tagged plants and merge the mapping and individual data:

```{r get-precise-locs, warning=FALSE}
vegmap <- getLocTOS(vegmap, "vst_mappingandtagging")

samples_to_keep <- vegind %>%
  filter(plantStatus == "Live") %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(individualID) %>%
  summarize(year = list(seq(min(year), max(year), by = 1))) %>%
  ungroup %>%
  unnest(year)

samples_to_keep %>%
  count(individualID) %>%
  arrange(-n)

veg <- as_tibble(vegmap) %>%
  group_by(individualID) %>%
  # some records are revised at later dates with better coordinates
  filter(!is.na(adjDecimalLatitude),
         date == max(date)) %>%
  ungroup %>%
  select(-date) %>%
  right_join(samples_to_keep) %>%
  distinct(individualID, year, .keep_all = TRUE)

veg %>%
  count(individualID, year) %>%
  arrange(-n)

veg %>%
  count(individualID) %>%
  arrange(-n)

veg %>%
  filter(individualID == "NEON.PLA.D03.JERC.02221") %>%
  View

nrow(vegmap)
nrow(veg)
```

Save output file:

```{r write-veg}
write_csv(veg, 'neon-veg.csv')
```

Visualize the imbalance of species identity:

```{r make-veg-tibble}
species_counts <- veg %>%
  count(scientificName) %>%
    separate(scientificName, into = c('genus', 'species'), sep = ' ',
           extra = 'drop', remove = FALSE) %>%
  mutate(name = paste(genus, species)) %>%
  arrange(-n) %>%
  mutate(scientific_name = fct_reorder(name, n),
         order = 1:n()) %>%
  dplyr::select(scientific_name, order, n)

species_counts %>%
  ggplot(aes(x = n, y = order)) +
  scale_x_log10() +
  geom_point(color = 'dodgerblue') +
  xlab('Number of mapped stem locations') +
  ylab('Order: commonest to rarest') +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
ggsave('~/Desktop/veg-freq.png', width = 4.5, height = 3)

# number of mapped stems
sum(species_counts$n)

# number of singletons
species_counts %>%
  filter(n == 1)

# number of species with fewer than 10 records
species_counts %>%
  filter(n < 10)

# number of species with fewer than 100 records
species_counts %>%
  filter(n < 100)
```

Split the veg data into different datasets for sites and year:

```{r split-veg}
split_veg <- veg %>%
  mutate(siteID_year = paste(siteID, year, sep = '_')) %>%
  split(.$siteID_year)
```

For each site-year combination, extract hyperspectral data for all plants:

```{r extract-data}
spectra <- vector(mode = 'list', length = length(split_veg))
dir.create('hs_data', showWarnings = FALSE)
csv_dir <- file.path("hs_data", "tabular")
dir.create(csv_dir, recursive = TRUE, showWarnings = FALSE)

try_extract <- function(path, pts) {
  out <- tryCatch(hs_extract_pts(path, pts, bands = 1:426),
                  error = function(e) NULL)
  return(out)
}

for (i in seq_along(split_veg)) {
  print(paste("processing year/site", i, "of", length(split_veg)))
  subd <- split_veg[[i]]
  identifier <- names(split_veg)[i]
  outfile <- file.path(csv_dir, paste0(identifier, ".csv"))
  if (!file.exists(outfile)) {
    h5_dir <- file.path('hs_data', identifier)
    dir.create(h5_dir, showWarnings = FALSE)
    siteyear <- identifier %>% str_split('_') %>% unlist
    site_code <- siteyear[1]
    year <- siteyear[2]

    # download data
    try(byTileAOP(dpID = "DP3.30006.001", site = site_code, year = year,
            easting = subd$adjEasting, northing = subd$adjNorthing,
            check.size = FALSE, savepath = h5_dir))
    hs_paths <- list.files(path = h5_dir, pattern = 'reflectance.h5',
                           recursive = TRUE, full.names = TRUE)
    if (length(hs_paths) > 0) {
      # extract spectra
      spdf <- SpatialPointsDataFrame(subd[, c('adjEasting', 'adjNorthing')],
                                 data = subd,
                                 proj4string = CRS(hs_proj4string(hs_paths[1])))
      out <- list()
      for (j in seq_along(hs_paths)) {
        res <- try_extract(hs_paths[j], spdf)
        first_band <- grep('^band1', names(res), value = TRUE)[1]
        if (!is.na(first_band)) {
          na_vals <- is.na(res[[first_band]])
          out[[i]] <- res[!na_vals, ]
        }
      }
      hs_df <- lapply(out, as.data.frame) %>%
        bind_rows %>%
        as_tibble %>%
        select(individualID, namedLocation,
               year, uid, namedLocation,
               adjEasting, adjNorthing,
               taxonID, taxonRank, scientificName,
               starts_with('band')) %>%
        distinct
      write_csv(hs_df, outfile)
      unlink(h5_dir, recursive = TRUE, force = TRUE)
    }
  }
}
```

Now, join all of the data together, and create one big data frame.

```{r join-df}
all_spectra <- list.files(csv_dir, full.names = TRUE) %>%
  lapply(function(f) {
    read_csv(f) %>%
      pivot_longer(starts_with("band")) %>%
      rename(reflectance = value) %>%
      separate(name, into = c("band_idx", "wavelength_nm")) %>%
      mutate(wavelength_nm = parse_number(wavelength_nm))
  })

nrows <- lapply(all_spectra, nrow) %>%
  unlist

all_spectra <- all_spectra[nrows > 0] %>%
  bind_rows %>%
  mutate(mask = case_when(
  .$wavelength_nm > 1340 & .$wavelength_nm < 1445 ~ TRUE,
  .$wavelength_nm > 1790 & .$wavelength_nm < 1955 ~ TRUE,
  .$wavelength_nm > 2400 ~ TRUE,
  TRUE ~ FALSE),
  spectraID = paste(individualID, year, sep = "_")) %>%
  select(spectraID, everything())

write_csv(all_spectra, "all_spectra.csv")
```

```{r plot-spectra}
all_spectra <- read_csv("all_spectra.csv")

p <- all_spectra %>%
  group_by(scientificName) %>%
  mutate(n = n()) %>%
  ggplot(aes(wavelength_nm, ifelse(mask, NA, reflectance), group = individualID)) +
  geom_path(alpha = .2) +
  xlab('Wavelength (nm)') +
  ylab('Reflectance') +
  facet_wrap(~reorder(scientificName, -n), labeller = label_wrap_gen()) +
  theme_minimal()
ggsave("spectra.png", plot = p, width = 15, height = 15)
```

Look at the sample sizes per species

```{r ss}
all_spectra %>%
  distinct(scientificName, spectraID) %>%
  count(scientificName) %>%
  ggplot(aes(x = n, y = reorder(scientificName, -n))) +
  geom_point() +
  ylab("Species") +
  xlab("Number of spectra") +
  theme(axis.text.y = element_text(size = 7)) +
  scale_x_log10()
```

Plot repeat samples of individuals:

```{r check-dups}
all_spectra %>%
  count(individualID) %>%
  arrange(-n) %>%
  top_n(1) %>%
  `[`(1:56, ) %>%
  left_join(all_spectra) %>%
  mutate(reflectance = ifelse(mask, NA, reflectance)) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, color = factor(year))) +
  geom_line() +
  facet_wrap(~individualID)
```