### 2020-09-10 R/clean_spectral_data.Rmd
added scripts for data cleaning and classifying

---
title: "Clean spectral data"
author: "Victoria Scholl"
date: "09/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(tidyverse)
library(kableExtra)
library(reshape2)
library(pbapply)
```

### Explore data

Read original spectra, extracted using NEON data (spectral-lib.Rmd).

Columns:

* **individualID** - unique identifier for each individual plant. Individuals can have multiple measurements over time.
* **uid** - unique identifier for the observation/record
* **spectraID** - unique identifier for each reflectance spectrum (a combination of individualID and year, YYYY )
* **namedLocation** - name of measurement location in NEON database
* **utmZone, adjEasting, adjNorthing** - UTM zone and coordinates of mapped stem
* **taxonID** - species code
* **scientificName** - scientific name, associated with taxonID, lowest taxonomic rank that can be determined
* **taxonRank** - lowest level taxonomic rank that can be determined for the individual (genus, species, variety)
* **band_idx** - number of the spectral band (band1, band 2, ... band 426)
* **wavelength_nm** - band wavelength in units of nm
* **reflectance** - decimal reflectance [0,1] extracted from NEON NIS hyperspectral imagery collected during same year
* **mask** - TRUE/LFALSE to identify water absorption bands

```{r}
# read_csv --> no factors, creates a tibble
all_spectra <- read_csv(here::here("analysis","data","raw_data",
                                   "spectral-lib","all_spectra.csv"))

# quick look at the values in each column
glimpse(all_spectra)
```

Number of unique **individualID**, **uid**, **spectraID** values:

```{r, message=TRUE}
message("# of unique individualID values:")
length(unique(all_spectra$individualID))

message("# of unique uid values:")
length(unique(all_spectra$uid))

message("# of unique spectraID (uid + year) values:")
length(unique(all_spectra$spectraID))
```

Number of unique **scientificName** & **taxonID** values (these should match):

```{r}
message("# of unique taxonID values:")
length(unique(all_spectra$taxonID))

message("# of unique scientificName values:")
length(unique(all_spectra$scientificName))
```

How many taxonomic identifications are at the genus vs species level? (Using the NEON-provided taxon Rank data)

```{r}
species_count <- all_spectra %>% 
  dplyr::distinct(spectraID, .keep_all = TRUE) %>%
  dplyr::group_by(taxonRank) %>% 
  dplyr::tally() %>% 
  kableExtra::kable() %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")

species_count
```

Filter the spectra to keep only those classified at the genus, species, and variety taxon ranks. Add new columns for genus and species by separating the **scientificName** field.

```{r}
# remove entries with taxonRank kingdom, family (maintain only genus, species, variety ranks)
all_spectra <- all_spectra %>% 
  dplyr::filter(taxonRank %in% c("genus", "species", "variety"))

# create new columns for genus and species (disregard variations within species) 
all_spectra <- all_spectra %>% 
  tidyr::separate(scientificName, into = c('genus', 'species'), sep = ' ', 
           extra = 'drop', remove = FALSE) %>% 
  # add a column combining genus and species for classification
  dplyr::mutate(genusSpecies = paste(genus,species))

message("# of unique genus values:")
length(unique(all_spectra$genus))

message("# of unique species values:")
length(unique(all_spectra$species))

message("# of unique genusSpecies values")
length(unique(all_spectra$genusSpecies))
```

```{r}
# example where there are multiple variations of a single genusSpecies value
all_spectra %>% 
  filter(genus == "Pseudotsuga") %>% 
  select(taxonID, genusSpecies, scientificName) %>% 
  distinct()
```

### Visualize spectra

Based on Max's Gist code: https://gist.github.com/mbjoseph/5c18781e508460e14f64193571b98b7d

Plot all spectra; color the spectra based on genus/species taxonomic level.

```{r, warning=FALSE, fig.height=16, fig.width=14}
all_spectra %>%
  ggplot(aes(wavelength_nm, ifelse(mask, NA, reflectance), group = spectraID,
             # color each facet plot based on taxonRank
             color = taxonRank)) +
  geom_path(alpha = .2) +
  labs(x = '\n Wavelength (nm)',
       y = 'Reflectance \n',
       color = "Taxon Rank") + 
  facet_wrap(~genusSpecies
             # wrap long names
             ,labeller = label_wrap_gen(20)) + 
  scale_color_manual(values = c("#e6550d","#1c9099", "#a8ddb5")) + 
  theme_minimal() + 
  # Adjust the text size and rotation angle 
  theme(text = element_text(size=11),
        axis.text.x = element_text(angle=90, hjust=1),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```

### Explore the bands

Look at the starting wavelength ("band1") for each reflectance spectrum:

```{r}
# how many different starting wavelength values in the data set? 
count_min_wavelengths <- all_spectra %>%
  dplyr::filter(band_idx == "band1") %>% 
  dplyr::group_by(wavelength_nm) %>%
  dplyr::count(wavelength_nm)

kableExtra::kable(count_min_wavelengths) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")
```

Filter spectra with starting wavelengths <380nm for now, since they complicate the wavelength interpolation.
All spectra with a starting wavelength of 347nm come from the [PUUM](https://www.neonscience.org/field-sites/field-sites-map/PUUM) site in Hawaii - why? TBD.

```{r}
# identify spectraID values that start at 347nm
spectra_to_remove <- all_spectra %>%
  dplyr::filter(band_idx == "band1") %>% 
  dplyr::filter(wavelength_nm == 347) %>% 
  dplyr::select(spectraID)

# remove these spectra from the data set
all_spectra <- all_spectra %>% 
  dplyr::filter(!(spectraID %in% spectra_to_remove$spectraID))

# how many different starting wavelength values in the filtered data set? 
count_min_wavelengths <- all_spectra %>%
  dplyr::filter(band_idx == "band1") %>% 
  dplyr::group_by(wavelength_nm) %>%
  dplyr::count(wavelength_nm)

kableExtra::kable(count_min_wavelengths) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")
```

Spectra have different starting wavelengths (i.e. 381, 383, 385nm...) and increase in increments of 5nm. Most of them start at 384nm.

```{r}
set.seed(44) #for obama

# find individualIDs that start with each of these wavelength band1 values
for(wl in unique(count_min_wavelengths$wavelength_nm)){

  tmp <- all_spectra %>% 
    dplyr::filter(band_idx == "band1") %>% 
    dplyr::filter(wavelength_nm == wl) %>% 
    dplyr::select(spectraID) %>% 
    dplyr::sample_n(1)

  message(paste("spectraID starting at", wl, "nm: ", as.character(tmp$spectraID)))

}
```

```{r warning=FALSE, fig.width=12, include=FALSE, eval=FALSE}
# plot spectra to see their starting/ending wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.4) +
  labs(title = "Spectra with different starting wavelengths") + 
  theme_bw()
```

Beginning of wavelength range:

```{r warning=FALSE, fig.width=12}
# plot spectra to see their starting wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra with different wavelengths: zoom in on beginning of spectra") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()
```

End of wavelength range:

```{r warning=FALSE, fig.width=12}
# plot spectra to see their ending wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra each with different wavelengths: zoom in on end of spectra") + 
  # zoom in on end of spectra
  lims(x = c(2455,2515), y = c(0, 0.3)) + 
  theme_bw()
```

### Standardize the bands: Interpolate wavelengths

Interpolate wavelength values to have the same 426 bands starting at 384 nm, ending at 2512 nm, incrementing by 5 nm. To avoid data gaps on the ends, let's use a larger range.

```{r wavelengths_interp}
# define the target wavelengths.
# use a larger range to get data on the ends.
# with a 5nm spacing, add 10nm to each end, and 4 to the total number of wavelengths. 
min_wl <- 384 
max_wl <- 2512 
num_wl <- 426
wavelength_vals <- seq(min_wl - 10, max_wl + 10,
                       length.out = num_wl + 4)
```

```{r}
# get bad bands for the target wavelengths

# identify a spectraID starting at the target wavelengths
tmp_spectrum <- all_spectra %>% 
    dplyr::filter(band_idx == "band1") %>% 
    dplyr::filter(wavelength_nm == min_wl) %>% 
    dplyr::select(spectraID) %>% 
    dplyr::sample_n(1)

# list the wavelengths that are bad bands, indicated by mask == TRUE
bad_bands <- all_spectra %>% 
  dplyr::filter(spectraID == tmp_spectrum$spectraID) %>%
  dplyr::filter(mask==TRUE) %>%
  dplyr::select(wavelength_nm, mask) 
```

```{r, warning=FALSE}
# interpolate all spectra efficiently 
interp_wl <- all_spectra %>%
  split(.$spectraID) %>%
  pblapply(function(df) {

    interpolated_reflectance <- approx(x= df$wavelength_nm, 
                                       y = df$reflectance, 
                                       xout = wavelength_vals)
    tibble(spectraID = unique(df$spectraID), 
           band_idx = seq_along(wavelength_vals), 
           # set wavelength values as integers
           wavelength_nm = round(interpolated_reflectance$x), 
           reflectance = interpolated_reflectance$y)

  }) %>%
  bind_rows %>% 
  # remove rows with wavelength values outside the range of interest 
  dplyr::filter(wavelength_nm >= min_wl) %>% 
  dplyr::filter(wavelength_nm <= max_wl) %>% 
  # update the band_idx column to be band1, ... band426
  dplyr::group_by(spectraID) %>% 
  dplyr::arrange(wavelength_nm) %>%
  dplyr::mutate(band_idx = paste0("band", 1:n())) %>%
  dplyr::arrange(spectraID)
```

```{r}
# set reflectance values to NA for bad bands 
interp_wl$reflectance[interp_wl$wavelength_nm %in% bad_bands$wavelength_nm] <- NA
```

Now, each spectrum (designated by a unique spectraID value) has the same 426 wavelength values. Reflectance values are set to NA for the bad band wavelength values. 
Merge the interpolated wavelengths and reflectance values into a copy of the data frame.

```{r, message=FALSE}
# make a copy of the original spectra 
cleaned_spectra <- all_spectra %>% 
  dplyr::select(-c(wavelength_nm, reflectance)) %>% 
  # merge in the interpolated wavelengths
  dplyr::left_join(interp_wl)
```

```{r, fig.width=12}
# plot interpolated spectra to see their starting wavelengths
interp_wl %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>% 
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra w interpolated wavelengths: Zoom in on beginning of spectra") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()
```

### Partition data into train, evaluation, & test sets

How many samples per species? Filter out species with just a few samples.

```{r filter_species_N_samples, fig.height=15, fig.width = 10}
# threshold to keep species with at more than this # of samples
thresh = 19

# count number of samples (unique spectra) per species 
taxonCounts <- cleaned_spectra %>% 
    distinct(spectraID, .keep_all = TRUE) %>% 
    #group_by(taxonID) %>% 
    group_by(genusSpecies) %>% 
    tally() %>% 
    # add a new column to indicate whether species count exceeds threshold
    mutate(countGT = case_when(
      n > thresh ~ TRUE,
      n <= thresh ~ FALSE
    ))

# count total number of samples across all species
samplesCountTotal <- taxonCounts %>% 
  summarise(sum(n))

speciesCountTotal <- nrow(taxonCounts)

# count total number of species with # samples greater than threshold
samplesCountGT <- taxonCounts %>% 
  filter(countGT == TRUE) %>% 
  summarise(sum(n))
speciesCountGT <- nrow(taxonCounts %>% 
  filter(countGT == TRUE))

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(taxonCounts, aes(y = reorder(genusSpecies, n), x = n)) + 
                          #reorder(taxonID, n), x = n)) + 
  geom_bar(stat = "identity", aes(fill = countGT)) + 
  labs(title = paste("Histogram of samples per species.",
                     speciesCountTotal, "total species;",
                     speciesCountGT, "species with greater than",
                     thresh, "samples"),
       y = "Genus Species",
       fill = paste("count greater than threshold", thresh)) + 
  theme_bw() + 
  scale_fill_manual(values = c("#d95f0e", "#7bccc4"))
```

```{r}
# Filter out species with less than N samples 
taxonCounts <- cleaned_spectra %>% 
    distinct(spectraID, .keep_all = TRUE) %>% 
    group_by(genusSpecies) %>% 
    tally() %>% 
  dplyr::filter(n>thresh)

filtered_spectra <- cleaned_spectra %>%
  dplyr::filter(genusSpecies %in% taxonCounts$genusSpecies)
```

Split multiple observations of individuals across sets. Add a column to the data (“train”, “eval”, “test”). The 80/20 ratio (based on the [Pareto Principle](https://en.wikipedia.org/wiki/Pareto_principle)) for training/testing is common, where the 20% can be split into 10/10 for testing/validation. Andrew NG recommends 60/20/20 for training/testing/validation in his online machine learning course.

```{r split_train_test_valid}
set.seed(44)
# efficiently split into train, test, valid groups 
splits <- filtered_spectra %>%
	distinct(spectraID) %>%
	mutate(group = sample(c("train", "valid", "test"), n(), 
	                      replace=TRUE, prob = c(.6, .2, .2)))

# check the percentage of each group
splits %>%
  group_by(group) %>% 
  tally() %>% 
  mutate(percent = n / sum(n))

# join these group labels into the spectra data frame 
filtered_spectra <- filtered_spectra %>%
	left_join(splits, by = "spectraID")
```

```{r}
# are all genusSpecies values present in each group?
filtered_spectra %>% 
  filter(group == "train") %>%
  select(genusSpecies) %>%
  unique() %>% 
  nrow()

filtered_spectra %>% 
  filter(group == "test") %>%
  select(genusSpecies) %>%
  unique() %>% 
  nrow()

filtered_spectra %>% 
  filter(group == "valid") %>%
  select(genusSpecies) %>%
  unique() %>% 
  nrow()
```

### Output "clean dataset"

Write spectra to csv file. 
```{r}
write_csv(x = filtered_spectra,
          path = here::here("data", "cleaned_spectra.csv"))
```

Visualize the spectra, facet by scientific name, color by taxon rank. 
```{r, warning=FALSE, fig.height=12, fig.width=16}
filtered_spectra %>%
  ggplot(aes(wavelength_nm, ifelse(mask, NA, reflectance), group = spectraID,
             # color each facet plot based on taxonRank
             color = taxonRank)) +
  geom_path(alpha = .2) +
  labs(x = '\n Wavelength (nm)',
       y = 'Reflectance \n',
       color = "Taxon Rank") + 
  facet_wrap(~genusSpecies
             # wrap long scientific names
             ,labeller = label_wrap_gen(20)
             ,ncol=10) + 
  scale_color_manual(values = c("#e6550d","#1c9099", "#a8ddb5")) + 
  theme_minimal() + 
  # Adjust the text size and rotation angle 
  theme(text = element_text(size=11),
        axis.text.x = element_text(angle=90, hjust=1),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```