### 2020-10-31 R/create_figures.Rmd
knits on laptop now

---
title: "Create additional figures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(dplyr)
library(sf)
library(ggplot2)
library(ggthemes)
library(spData)
library(tidyverse)
library(RColorBrewer)
library(ggspatial)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```

## Study area

```{r prepare_data, message=FALSE, warning=FALSE}

# download and unzip NEON domain data
url <- "https://www.neonscience.org/sites/default/files/NEONDomains_0.zip" 
dest_dir <- here::here("analysis", "data", "shp_data")
if(!dir.exists(dest_dir)){
  dir.create(dest_dir)
}
dest_file <- here::here("analysis", "data", "shp_data", "NEONDomains.zip")
download.file(url, dest_file)
unzip(dest_file, exdir = dest_dir)

# read NEON domain polygons
neon_domains <- sf::st_read(here::here("analysis", "data", "shp_data", "NEON_Domains.shp"), 
                            quiet = TRUE)

# read spectral data
cleaned_spectra <- readr::read_csv(here::here("analysis", "data", "derived_data", 
                                              "cleaned_spectra.csv"))
spectra_locations <- cleaned_spectra %>%  
  # grab the latitude and longitude coordinates for each spectrum 
  select(latitude = adjDecimalLatitude, longitude = adjDecimalLongitude) %>% 
  # remove redundant rows
  distinct() %>%
  # convert each to a SF point with geographic coordinates 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# filter the NEON domains, keep only those containing spectra locations
domains_containing_spectra <- neon_domains %>%
  filter(lengths(st_intersects(neon_domains, spectra_locations)) > 0)

# transform domains and spectra points to Albers Equal Area, CRS = 2163 for the map 
domains_filtered_2163 <- sf::st_transform(x = domains_containing_spectra, crs = 2163)
domains_all_2163 <- sf::st_transform(x = neon_domains, crs = 2163)
spectra_2163 <- sf::st_transform(x = spectra_locations, crs = 2163)

# load and transform additional geospatial data for mapping 
data("us_states", package = "spData")
us_states_2163 <- sf::st_transform(spData::us_states, crs = 2163)
#world_2163 <- st_transform(spData::world, crs = 2163)
world_2163 <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>% st_transform(crs = 2163)

# set color palette for NEON domains 
colourCount <- length(unique(domains_filtered_2163$DomainID))

# colorbrewer palette can only be up to 11 colors though
getPalette <- colorRampPalette(brewer.pal(colourCount, "Spectral"))
colors_domains <- data.frame(names = sort(unique(domains_filtered_2163$DomainName)),
                        values = getPalette(colourCount)) #colorRampPalette

# trying to use a non-color brewer palette to get 14
#getPalette <- c(brewer.pal(name="Spectral", n = 8), brewer.pal(name="Spectral", n = colourCount-8))
#colors_domains <- data.frame(names = sort(unique(domains_filtered_2163$DomainName)),
#                        values = getPalette) 
```

```{r make_map, fig.height=4, fig.width=6}
# make the map 
ggplot() + 
  # US States
  geom_sf(data = us_states_2163, fill="grey80", color="gray70", alpha=0.3) + 
  # continent boundaries for context 
  geom_sf(data = world_2163, fill="grey80", color="gray50", alpha=0.5) +
  # NEON domains containing spectra
  geom_sf(data = domains_filtered_2163, aes(fill = factor(DomainName)), 
          alpha = 0.7, color = "gray40") + 
  # fill label in legend 
  labs(fill = "NEON Domain") + 
  # apply fill color scale for neon domains 
  scale_fill_manual(breaks = colors_domains$names, values = colors_domains$values) + 
  # add the other NEON domains, just as outlines + 
  geom_sf(data = domains_all_2163, alpha = 0, color = "gray40") +
  # add NEON spectra locations as points
  geom_sf(data = spectra_2163) + 
  # set the extent based on NEON domains where we have spectral data 
  coord_sf(crs = st_crs(2163),
           expand = FALSE,
           xlim = c(as.numeric(st_bbox(domains_filtered_2163)[1]),    # xmin
                    as.numeric(st_bbox(domains_filtered_2163)[3])),   # xmax 
           ylim = c(as.numeric(st_bbox(domains_filtered_2163)[2]),    # ymin
                    as.numeric(st_bbox(domains_filtered_2163)[4]))) + # ymax
  # add scale bar and north arrow 
  ggspatial::annotation_scale(location = "bl", width_hint = 0.4, style = "bar") +
  # add north arrow
  ggspatial::annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +
  # set theme 
  theme_bw() 
```

```{r make_map_with_alaska_inset, fig.height=4, fig.width=6}
# make the map WITH ALASKA INSET
main_map <- ggplot() + 
  # US States
  geom_sf(data = us_states_2163, fill="grey80", color="gray70", alpha=0.3) + 
  # continent boundaries for context 
  geom_sf(data = world_2163, fill="grey80", color="gray50", alpha=0.5) +
  # NEON domains containing spectra
  geom_sf(data = domains_filtered_2163, aes(fill = factor(DomainName)), 
          alpha = 0.7, color = "gray40") + 
  # fill label in legend 
  labs(fill = "NEON Domain") + 
  # apply fill color scale for neon domains 
  scale_fill_manual(breaks = colors_domains$names, values = colors_domains$values) + 
  # add the other NEON domains, just as outlines + 
  geom_sf(data = domains_all_2163, alpha = 0, color = "gray40") +
  # add NEON spectra locations as points
  geom_sf(data = spectra_2163) + 
  # set the extent based on NEON domains where we have spectral data 
    coord_sf(crs = st_crs(2163), 
           xlim = c(-2500000, 2500000), 
           ylim = c(-2300000, 730000))  + 
  # get rid of legend
  theme(legend.position = "none")

alaska_inset <- ggplot() + 
  # add Alaska boundaries
  geom_sf(data = us_states_2163, fill="grey", color="gray70", alpha=0.3) + 
  # continent boundaries for context 
  geom_sf(data = world_2163, fill="grey", color="gray70", alpha=0.3) + 
  # Add NEON domains 
  geom_sf(data = domains_filtered_2163, aes(fill = factor(DomainName)), 
          alpha = 0.7, color = "gray40") + 
  # add NEON sites as points 
  geom_sf(data = spectra_2163) + 
  # set extents to zoom in just on Alaska
  coord_sf(crs = st_crs(3467), 
           xlim = c(-2400000, 1600000), 
           ylim = c(200000, 2500000), 
           expand = FALSE, datum = NA) + 
  theme_clean() +
  # get rid of legend for now
  theme(legend.position = "none") + 
  # fill for neon domains in the legend 
  scale_fill_manual(breaks = colors_domains$names, values = colors_domains$values) + 
  # add a border around the alaska inset 
  theme(panel.border = element_rect(colour = "black", fill = NA))

# combine main map with alaska as an inset
main_map +
  annotation_custom(grob = ggplotGrob(alaska_inset),
                    xmin = -2750000,
                    xmax = -2700000 + (1600000 - (-2400000))/2.5,
                    ymin = -2470000,
                    ymax = -2450000 + (2500000 - 20000)/2.5) + 
  # set theme to make background white, subtle grid lines, outline map
  theme_bw()  + 
  # apply color scale for domains across the contiguous US and Alaska
  scale_fill_manual(breaks = colors_domains$names, values = colors_domains$values) + 
  # titles and labels
  labs(fill = "NEON Domain") + 
  # add scale bar
  annotation_scale(location = "br", width_hint = 0.5, style = "bar") +
  # add north arrow
  annotation_north_arrow(location = "tl", which_north = "true", 
                         pad_x = unit(0.15, "in"), pad_y = unit(0.5, "in"),
                         style = north_arrow_fancy_orienteering) 
```

```{r}
message("saving study area figure...")
ggsave(filename = here::here("analysis", "figures", "study_area.pdf"),
       width = 12, height = 8)
```

## Workflow figure

```{r include=FALSE, eval=FALSE}
library(DiagrammeR)
library(rsvg)

workflow <- DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

data1 [label = 'In-situ plant \nmeasurements', shape = folder, fillcolor = LightSteelBlue3]
data2 [label = 'Hyperspectral \nreflectance', shape = folder, fillcolor = LightSteelBlue3]
process1 [label =  'Extract \n Reflectance', fillcolor = Gainsboro]
process2 [label = 'Data cleaning & \nPreparation', fillcolor = Gainsboro]
classify1 [label = 'Classify Species', fillcolor = AntiqueWhite3]
classify2[label = 'Classify Genus', fillcolor = AntiqueWhite3]
results [label= 'Results', shape = oval, fillcolor = Honeydew3]

# edge definitions with the node IDs
{data1 data2}  -> process1 -> process2 -> {classify1 classify2} -> results
}")

workflow

export_graph(workflow,
  file_name = here::here("analysis", "figures", "workflow.pdf"),
  file_type = "pdf")

DiagrammeR::export_graph(workflow) %>%
  charToRaw() %>%
  rsvg::rsvg_pdf(here::here("analysis", "figures", "workflow.pdf"))

#grViz(workflow) %>%
#    export_svg %>% charToRaw %>% rsvg_pdf(here::here("analysis", "figures", "workflow.pdf"))
```