### 2020-11-10 analysis/paper/paper.Rmd
working on intro text

---
title: \textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

author:
  - Victoria Scholl\thanks{victoria.scholl@colorado.edu, Earth Lab and Geography Dept., University of Colorado, Boulder, CO 80303, USA}
  - Maxwell B. Joseph\thanks{maxwell.b.joseph@colorado.edu, Earth Lab, University of Colorado, Boulder, CO 80303, USA}

output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    includes:
      in_header: header.sty
bibliography: references.bib
csl: ecology.csl 
---

```{r setup, echo = FALSE}
library(bookdown)
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  comment = "#>",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

\linenumbers

\begin{abstract}
Abstract text here
\end{abstract}

# Introduction {-}

Mapping forests is crucial to understanding the effects and feedbacks of changes in climate, land use, and disturbance events across diverse landscapes [@saladin2020rapid; @grossiord2020having; @fassnacht2016review]. 
An important aspect of forest mapping efforts is species composition, which describes the amount and distribution of plant species across landscapes. 
Species composition has been shown to be a fundamental metric for evaluating biodiversity [@williams2020remote], identifying habitats [@nagendra2013remote], providing essential inputs to many ecological models [@pecchi2019species], and informing conservation and restoration efforts [@camarretta2020monitoring]. 
Although remote sensing technologies enable us to estimate species composition more effectively than using field-based approaches alone, species classification is still a challenging task.

We currently have unprecedented access to publicly available hyperspectral airborne imagery at sites around the United States from the National Ecological Observatory Network (NEON) [@keller2008continental]. 
NEON-monitored sites are spread across ecoclimatic domains with a diverse array of ecosystems and land cover types including deciduous and evergreen forests, shrub/scrub and woody wetlands, mixed forest and grasslands. 
Field technicians collect field-based vegetation measurements including species observations on a semi-annual basis. 
In addition, NEON’s Airborne Observation Platform (AOP) payload carries a suite of sensors to collect both spectral and structural measurements of ecosystems [@kampe2010advances]. 
The AOP includes an imaging spectrometer which captures hyperspectral reflectance data in over 400 bands across the visible and near infrared wavelengths. 
Hyperspectral imagery is frequently used to classify species with accuracies ranging widely anywhere from 45% to 98% [@fassnacht2016review]. 
Hyperspectral data are capable of capturing reflectance variations and features with hundreds of spectral bands better than multispectral remotely sensed data with a handful to tens of bands [@awad2018forest].
Many factors influence tree species classification results including the number of species being discerned, the homogeneity of species, the density and structural complexity of the forest, and the taxonomic level of classification [@sothe2019tree]. 
Methods are often performed on small scales or in simple ecosystems, and are not successful at scaling to larger areas or different forest types **(cite)**.
Although the demand for species mapping is justified in many cases, there is also potential for both using and finding value in classifying individuals at other taxonomic levels, as explored in a recent bioacoustics study [@cramer2020chirping]. 
In the plant realm, identifying taxons at the genus level had been shown to be easier than at the species level [@tuominen2018assessment].
[@surovy2019acquisition] report that species are more successfully classified into broader categories such as spruce, pine, and broadleaf, compared to identifying individual species in mixed arrangements.

NEON provides a unique opportunity to test, develop, and tune tree classification methods over a continental scale and at multiple taxonomic resolutions. 
We have extracted all NEON AOP hyperspectral data for every mapped stem in NEON plots where field data and airborne remote sensing data were collected during the same year.
We used a machine learning approach to evaluate plant identification potential using the hyperspectral data at multiple taxonomic resolutions: species and genus levels. 
Our study contributes to the active field of tree species classification research using freely available NEON data and open, reproducible code [@marconi2019data; @scholl2020integrating]

# Background {-}

# Methods {-}

## NEON data {-}

The NEON project monitors diverse ecosystems at 81 sites across 20 domains, each with similar climate and ecological characteristics, across the United States.
The NEON data that we used for our study include both *in-situ* plant measurements and airborne remote sensing image data [@neonDataProducts].
The Woody Plant Vegetation Structure (data product code DP1.10098) includes individual plant locations along with species and genus observations that are collected every 1 to 3 years. 
The NEON Airborne Observation Platform (AOP) collects airborne remote sensing data regularly at NEON sites. 
There are three AOP payloads, each carrying three types of sensors: (1) a high-resolution digital camera to capture red, green, blue (RGB) true-color images, (2) a light detection and ranging (lidar) system to capture discrete and waveform lidar data with approximately 1-4 points/waveforms per square meter, and (3) a pushbroom-style AVIRIS next-gen NEON Imaging Spectrometer (NIS) to capture hyperspectral data. 
The hyperspectral data contains 426 narrow bands spanning the visible to shortwave infrared wavelengths, from 380 - 2500nm, and the image data products derived have a spatial resolution of 1m. The AOP flies at an altitude of 1000m above ground level and the flight season typically runs from May to October. 
We worked with the NIS-derived Spectrometer orthorectified surface directional reflectance mosaic (data product code DP3.30006).

## Data cleaning and preparation {-}

We performed our analysis in R [@R] with packages including neonUtilities [@neonUtilities], geoNEON [@geoNEON], neonhs [@neonhs].
In June 2020, we extracted hyperspectral reflectance data for all individual mapped stems at NEON sites that had an AOP flyover during the same year as when *in-situ* measurements were collected. 
We only kept individuals with a "live" plant status (as opposed to other plant status classifications such as "dead", "standing dead", "downed", "insect/disease damaged", and "no longer qualifies").
This included spectral reflectance spectra for 5972 observations of individual plants. 
Some individuals had multiple observations across years. 
The individual plants were identified taxon ranks of kingdom, family, genus, species, and variety. 
Most of them were identified at the species rank.
We filtered the plants by taxonomic rank to retain only those identified at the genus, species, and variety ranks. 
We removed varieties from scientific names to label each individual spectrum with its respective genus and species (when available), represented by the first 2 words in the scientific name. 
There were 74 unique genus values, 113 unique species values, and 151 unique genus-species combinations.
The hyperspectral reflectance spectra extracted from the locations of the individual plants have different starting wavelengths:

```{r wavelengthTable, message = FALSE, echo = FALSE}
df <- data.frame(wavelength = c(347, 381, 382, 384),
                 number = c(94, 1220, 1733, 2922))

knitr::kable(x = df, 
             # set column names
             col.names = c("Starting wavelength [nm]",
                           "Number of spectra"),
             # center align text in the columns
             align = c("c", "c"),
             format = "simple")
```

Each series of wavelengths increments by 5nm from each starting wavelength to yield 426 spectral bands. 
These starting wavelengths vary depending on which AOP payload was used to collect the hyperspectral data.
The wavelength values also change slightly each year when the NEON AOP re-calibrates their payload sensors [@kampe2010advances]. 
Most of them started at 384nm, so we linearly interpolated each spectrum to have the same 426 bands starting at 384nm and ending at 2512nm with 5nm increments.
Spectra from the Pu'u Maka'ala Natural Area Reserve (PUUM) NEON site in Hawaii had a starting wavelength of 347nm.
This is because NEON sub-contracted the PUUM flights to the Global Airborne Observatory, which has a spectrometer with a different wavelength range compared to the NIS payloads *([how to cite?](https://www.neonscience.org/observatory/observatory-blog/remote-sensing-partnership-made-paradise))*. 
We filtered out any spectra with a starting wavelength of 347nm, since the interpolation would require too much unavailable data outside the available range.
We flagged "bad bands" from 1340-1445nm, 1790-1995nm, and greater than 2400nm so they could be excluded from later analysis.
These wavelength ranges feature noisy reflectance data due to high water absorption [@bajorski2011].

Next, we removed samples belonging to species and genus classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 151 unique scientific names identified at the genus or species level, 50 of them had at least 20 samples and were kept for the classification analysis (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 scientific name labels were represented in each set.

\clearpage

```{r taxonTable, message = FALSE, echo = FALSE}
library(knitr)
library(dplyr)

df <- data.frame(taxonCode = c("ABBA", "ABLAL", "ACRU, ACRUR", "ACSAS, ACSA3", 
                               "AMLA", "ARTR2", "BELE", "BENE4", "CACA18", "CAOV2",
                               "CATO6", "CELA", "CODR", "FRAM2", "FRPE", "JUNI", 
                               "JUVIV, JUVI", "LARIX", "LATR2", "LIST2", "LITU",
                               "LOMA6", "MORU2", "NYBI", "NYSY", "PIEN", "PIMA",
                               "PICOL, PICO", "PIEC2", "PIFL2", "PIPA2", "PITA",
                               "POGR4", "POTR5", "PSMEM, PSME", "QUAL", "QUCO2",
                               "QUHE2", "QULA2", "QUMA13", "QUMA3", "QUNI", "QUPA5",
                               "QURU", "QUST", "RHDA", "TSCA", "TSHE", "ULAM", "ULCR"),
                 scientificName = c("*Abies balsamea*", "*Abies lasiocarpa*", "*Acer rubrum*",
                                    "*Acer saccharum*", "*Amelanchier laevis*", "*Artemisia tridentata*",
                                    "*Betula lenta*", "*Betula neoalaskana*", "*Carpinus caroliniana*",
                                    "*Carya ovata*", "*Carya tomentosa*", "*Celtis laevigata*", 
                                    "*Cornus drummondii*", "*Fraxinus americana*", 
                                    "*Fraxinus pennsylvanica*", "*Juglans nigra*", 
                                    "*Juniperus virginiana*", "*Larix sp.* *",
                                    "*Larrea tridentata*", "*Liquidambar styraciflua*", 
                                    "*Liriodendron tulipifera*", "*Lonicera maackii*",
                                    "*Morus rubra*", "*Nyssa biflora*", "*Nyssa sylvatica*",
                                    "*Picea engelmannii*", "*Picea mariana*", "*Pinus contorta*",
                                    "*Pinus echinata*", "*Pinus flexilis*","*Pinus palustris*", 
                                    "*Pinus taeda*", "*Populus grandidentata*", "*Populus tremuloides*",
                                    "*Pseudotsuga menziesii*", "*Quercus alba*", "*Quercus coccinea*",
                                    "*Quercus hemisphaerica*", "*Quercus laevis*", "*Quercus margaretta*",
                                    "*Quercus marilandica*", "*Quercus nigra*", "*Quercus pagoda*",
                                    "*Quercus rubra*", "*Quercus stellata*", "*Rhamnus davurica*",
                                    "*Tsuga canadensis*", "*Tsuga heterophylla*", "*Ulmus americana*", 
                                    "*Ulmus crassifolia*"),
                 commonName = c("balsam fir", "subalpine fir", "red maple", "sugar maple",
                                "Allegheny serviceberry", "sagebrush", "sweet birch", "resin birch",
                                "American hornbeam", "shagbark hickory", "mockernut hickory",
                                "sugarberry", "roughleaf dogwood", "white ash", "green ash",
                                "black walnut", "eastern redcedar", "larch", "creosote bush",
                                "sweetgum", "tuliptree", "Amur honeysuckle", "red mulberry",
                                "swamp tupelo", "blackgum", "Engelmann spruce", "black spruce",
                                "lodgepole pine", "shortleaf pine", "limber pine", "longleaf pine",
                                "loblolly pine", "bigtooth aspen", "quaking aspen", "Douglas-fir",
                                "white oak", "scarlet oak", "Darlington oak", "turkey oak",
                                "sand post oak", "blackjack oak", "water oak", "cherrybark oak",
                                "northern red oak", "post oak", "Dahurian buckthorn", 
                                "eastern hemlock", "western hemlock", "American elm", "cedar elm"),
                 count = c(47,117, 489, 79, 37, 494, 28, 35, 170, 46, 134, 179, 31, 20, 92, 42, 
                           28, 23, 161, 209, 40, 23, 30, 24, 77, 24, 32, 302, 72, 78, 374, 87,
                           54, 148, 88, 72, 29, 69, 132, 68, 53, 60, 30, 148, 151, 224, 161, 59,
                           70, 26))

df$count_percent <- round((df$count / sum(df$count)) * 100, digits = 1)

knitr::kable(x = df, 
             # set column names
             col.names = c("Taxon code",
                           "Scientific name (*Genus species*)",
                           "Common name",
                           "Number of spectra",
                           "Percent (%) of total"),
             # center align text in the columns
             align = c("l", "l", "l", "c", "c"),
             format = "simple") 
  # show which rows are grouped together for genus classes visibly in table?

```

After these data cleaning steps, we were left with reflectance data for 5266 individuals at sites across 14 of the NEON eco-climatic domains (Figure \@ref(fig:fig-studyArea)). 
There is a noticeable imbalance of data across space and time; some NEON domains and US states (such as Hawaii) are not represented in our analysis. 
Years that we found when both *in-situ* and AOP data were collected were 2017, 2018, and 2019. 
Note that covid-19 **(cite?)** postponed NEON data collection as well as the AOP flight season during 2020. 
We assumed that plant locations did not change over time.

Show the spectra for each scientific name class (Figure \@ref(fig:fig-plotSpectra)).

## Species and genus classification {-}

We trained a random forest (RF) classifier to predict species based on NEON hyperspectral reflectance data as the descriptive features.
RF is a machine learning approach that utilizes an ensemble of decision trees and iterative subsets of the training data.
We performed hyperparameter optimization of *ntree*, the number of decision trees in the forest, and *mtry*, the number of variables randomly sampled at each node in the tree.
We used the optimal values of 27 for *mtry* and 1000 for *ntree* in our final RF model.
There were 49 species classes and 27 genus classes for each classification. 
We computed overall accuracy, confusion matrices, and variable importance plots.

To evaluate genus classification accuracy, we followed the same step as described for species classification, but we used each spectrum's associated genus as each class label. 
We used the randomForest R package [@randomForest] to implement the analysis. 
We provide all of our code in a research compendium here:[https://github.com/vscholl/neonHScompendium](https://github.com/vscholl/neonHScompendium).

# Results {-}

We report overall classification accuracy of the species and genus classification:

```{r, message = FALSE, echo = FALSE}
library(knitr)

df <- data.frame(oa = c("Out-of-bag (OOB)", "Independent validation"),
                 species = c(54.6, 54.9),
                 genus = c(60.5, 60.0))

knitr::kable(x = df, 
             # set column names
             col.names = c("Overall Classification Accuracy [%]",
                           "Species",
                           "Genus"),
             # center align text in the columns
             align = c("c", "c", "c"),
             format = "simple")
```

Classification accuracy was higher at the genus level (60%) compared to the species level (55%) based on both the internal OOB error of the RF model and the independent validation set.

Confusion matrix for species classification (Figure \@ref(fig:fig-confMatSpecies)).

Confusion matrix for genus classification (Figure \@ref(fig:fig-confMatGenus)).

Variable importance for species and genus classifiers (Figure \@ref(fig:fig-varImp)).

# Discussion {-}

Restate findings at high level

How do our findings compare to other similar studies. Consistent? Differences? (cite IDTReeS) Other ranges of accuracy using NEON HS data? How many sites too?

Potential limitations
  - class imbalance - we tried to filter based on # samples, but our model has then only seen common species and would be unable to identify rare ones
  - filtering shadows, non-vegetation pixels.... (cite a paper from last IDTReeS comp)

Future directions, next steps
  - targeted species/genus for increased sampling to improve representation in the data set?
   - more advanced models. ML DL
   - RF treats each feature as independent
   - spectra are sequences of wavelengths - explore DL methods for sequences (i.e. recurrent neural networks)
   - hierarchical taxonomic approach - borrow information from closely related species (i.e. common and rare species > use common species observations)
   - (cite our other paper) intelligent crown geometry processing 
   - (ben weinstein) tree crown detection - extract multiple pixels per individual using rectangular crown boundaries
   - herbarium / hyperspectral data measurements at other scales. handheld spectrometer. other spectral libraries.
   - combine HS with RGB, lidar

We conclude that NEON hyperspectral data allow you to achieve 55% or 60% accuracy when classifying species or genus classes.

Hopes for outcomes

# Appendix / Supplement ideas {-}

- Species histogram to show which ones were excluded for few # samples 
- Wavelength interpolation figures?

# Acknowledgements {-}

Max has cool ideas and knows many R tricks

\clearpage

# References {-}

<div id="refs"></div>

\newpage

\clearpage

# Figure legends {-}

## Figure 1 {-}

Map to illustrate where NEON data were collected that we used in this study. Black points represent NEON site locations where *in-situ* plant observations were collected along with airborne remote sensing flights. NEON domains containing each respective site are

## Figure 2 {-}

Histogram showing which scientific names have at least 20 samples to be included in the classification analysis. Note that the "Larix sp." individuals are classified at the genus level.

## Figure 3 {-}

Visualizing the extracted hyperspectral reflectance spectra for each of the 50 genus and species classes that we used in our classifications.

## Figure 4 {-}

Confusion matrix for the species classifier. True and predicted species labels are shown here for each sample in the validation set. Counts along the diagonal represent correct species predictions. Counts outside of the diagonal represent incorrect species predictions. Each cell in the matrix is colored based on the number of counts to ease interpretation. Low counts are white and high counts are blue, with a gradient in between. Counts of zero are shown with a gray font color, to make non-zero counts (shown with a black font color) easier to identify.

## Figure 5 {-}

Confusion matrix for the genus classifier. True and predicted genus labels are shown here for each sample in the validation set. Counts along the diagonal represent correct genus predictions. Counts outside of the diagonal represent incorrect genus predictions. Each cell in the matrix is colored based on the number of counts to ease interpretation. Low counts are white and high counts are orange, with a gradient in between. Counts of zero are shown with a gray font color, to make non-zero counts (shown with a black font color) easier to identify.

## Figure 6 {-}

Variable importance plots for the species and genus classification. We used hyperspectral reflectance band values as descriptive features for each classifier. The Mean Decrease in Accuracy (MDA) importance metric value is plotted for each hyperspectral band. Each importance bar is colored based on importance, with low importance as white and follows a gradient to high importance as blue (for species variable importance) or orange (for genus variable importance).

# Figures {-}

## Figure 1 {-}

```{r fig-studyArea, fig.cap ="Map", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis", "figures", "study_area.pdf"))
```

\clearpage

## Figure 2 {-}

```{r fig-histSpecies, fig.cap ="Histogram", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "hist_samples_per_species.pdf"))
```

\clearpage

## Figure 3 {-}

```{r fig-plotSpectra, fig.cap ="Spectra", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "spectra_per_species.pdf"))
```

\clearpage

## Figure 4 {-}

```{r fig-confMatSpecies, fig.cap ="Confusion matrix for species classification", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "confMat_species_valid.pdf"))
```

\clearpage

## Figure 5 {-}

```{r fig-confMatGenus, fig.cap ="Confusion matrix for genus classification", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "confMat_genus_valid.pdf"))
```

\clearpage

## Figure 6 {-}

```{r fig-varImp, fig.cap ="Variable importance", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "varImp.pdf"))
```

\clearpage