### 2021-04-22 R/add_plant_functional_type.Rmd
up

---
title: "Add Plant Functional Type (PFT) labels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE}
library(tidyverse)
library(here)
library(dplyr)
library(kableExtra)
library(neonUtilities)
```

Plant functional types (PFTs) originally proposed: coniferous forest, deciduous forest, woody shrub, grass/herb, and bare ground.

Proposing to update the PFTs based on the MODIS land cover types: Evergreen Needleleaf
Forests, Evergreen Broadleaf Forests, Deciduous Needleleaf Forests, Deciduous Broadleaf
Forests, Shrub.

```{r load_cleaned_spectra, message=FALSE}
# start timer 
start_time <- Sys.time()

# read cleaned spectral library file 
cleaned_spectra <- read_csv(here::here("analysis", "data", "derived_data", "cleaned_spectra.csv"))
```

## Use NEON growthForm observations

Run this code chunk to download the NEON woody vegetation structure data: 
```{r download_neon_veg_data, warning=FALSE, message=FALSE, eval=FALSE}
# download woody vegetation structure data for all sites and years 
out_dir <- here::here("analysis","data","raw_data")

neonUtilities::zipsByProduct(dpID = "DP1.10098.001",
              savepath = out_dir,
              check.size = FALSE)

stackByTable(file.path(out_dir, "filesToStack10098"), folder = TRUE)
```

```{r read_merge_simplify_growthforms}
# read the vst_apparentindividual table, which has growthForm data
vst_apparentindividual <- read_csv(out_dir <- here::here("analysis","data","raw_data", "filesToStack10098",
                                                         "stackedFiles", "vst_apparentindividual.csv"), 
                                   col_types = cols())

# select the growthForm and height columns
growthForms <- vst_apparentindividual %>% 
  select(individualID, growthForm, height)

# join the growthForm and height columns into the spectra data frame
cleaned_spectra_merged <- right_join(cleaned_spectra, growthForms, by = "individualID")

# print all unique growthForm values 
unique(cleaned_spectra_merged$growthForm)

# simplify the growth form into "tree", "shrub", "sapling" etc.  
cleaned_spectra_merged <- cleaned_spectra_merged %>% 
  rowwise() %>%
  # parse out the last word in each entry's growthForm
  # make this code more efficient????? 
  mutate(growthForm_simple = rev(strsplit(growthForm, " ")[[1]])[1])
```

```{r summary_table_taxon_growthforms}
# summarize the merged growthform data
cleaned_spectra_merged %>% 
  distinct(spectraID, .keep_all = TRUE) %>% 
  group_by(genusSpecies, growthForm_simple) %>% 
    tally()

# summarize the merged growthform data with "tree" and "shrub" growth forms
cleaned_spectra_merged %>% 
  distinct(spectraID, .keep_all = TRUE) %>% 
  # keep just the tree and shrub growth forms, remove rows with "sampling" and "liana"
  filter(growthForm_simple %in% c("tree","shrub")) %>% 
  group_by(genusSpecies, growthForm_simple) %>% 
    tally() %>% 
  # display as a nicely formatted table
  kableExtra::kable() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            full_width = F, 
                            position = "left")

```

## Manually add PFTs to spectral library

This code manually adds labels to each taxon (genus and species listed here) for:

- common name
- PFT (based on Macrosystems labels)
- seasonality (evergreen or deciduous)
- leaf type (broadleaf or needleleaf)
- growth form (shrub or forest)

```{r manually_add_pft_labels}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
# PFTs of interest: coniferous_forest, deciduous_forest, woody_shrub, grass_herb, and bare_ground.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft = NA,
                               seasonality = NA,
                               leafType = NA,
                               growthForm = NA) %>% 
  # arrange alphabetically
  dplyr::arrange(genusSpecies)

for(i in 1:nrow(taxon_pft_lookup)){
  # print row and current genusSpecies label
  #print(i)
  genusSpecies <- taxon_pft_lookup$genusSpecies[i]
  #print(genusSpecies)

  if(genusSpecies == "Abies amabilis"){
    taxon_pft_lookup$pft[i] = "coniferous_forest" 
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "Pacific silver fir"}

  else if(genusSpecies == "Abies balsamea"){
    taxon_pft_lookup$pft[i] = "coniferous_forest" 
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "balsam fir"}

  else if(genusSpecies == "Abies lasiocarpa"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "needleleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "subalpine fir"}

  else if(genusSpecies == "Acacia greggii"){
    taxon_pft_lookup$pft[i] = "woody_shrub" 
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "catclaw acacia"}

  else if(genusSpecies == "Acer barbatum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "Southern sugar maple"}

  else if(genusSpecies == "Acer circinatum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest" 
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "vine maple"}

  else if(genusSpecies == "Acer negundo"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "boxelder"}

  else if(genusSpecies == "Acer pensylvanicum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "striped maple"}

  else if(genusSpecies == "Acer rubrum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "red maple"}

  else if(genusSpecies == "Acer saccharinum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "silver maple"}

  else if(genusSpecies == "Acer saccharum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "sugar maple"}

  else if(genusSpecies == "Ailanthus altissima"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "tree of heaven"}

  else if(genusSpecies == "Alnus rubra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "red alder"}

  else if(genusSpecies == "Aloysia wrightii"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "Wright's beebrush"}

  else if(genusSpecies == "Amelanchier laevis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "Allegheny serviceberry"}

  else if(genusSpecies == "Amyris elemifera"){
    taxon_pft_lookup$pft[i] = "???" # is not deciduous or coniferous forest
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "sea torchwood"}

  else if(genusSpecies == "Artemisia tridentata"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "big sagebrush"}

  else if(genusSpecies == "Atriplex canescens"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "fourwing saltbush"}

  else if(genusSpecies == "Atriplex confertifolia"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$seasonality[i] = "evergreen"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "shrub"
    taxon_pft_lookup$commonName[i] = "shadscale saltbush"}

  else if(genusSpecies == "Betula alleghaniensis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "yellow birch"}

  else if(genusSpecies == "Betula lenta"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "sweet birch"}

  else if(genusSpecies == "Betula neoalaskana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "resin birch"}

  else if(genusSpecies == "Betula nigra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "river birch"}

  else if(genusSpecies == "Betula papyrifera"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "paper birch"}

  else if(genusSpecies == "Betula sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = "broadleaf"
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "birch"}

  else if(genusSpecies == "Bourreria sp."){
    taxon_pft_lookup$pft[i] = "???"   
    taxon_pft_lookup$seasonality[i] = ""
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "strongbark"}

  else if(genusSpecies == "Bursera simaruba"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"  
    taxon_pft_lookup$seasonality[i] = "deciduous"
    taxon_pft_lookup$leafType[i] = ""
    taxon_pft_lookup$growthForm[i] = "forest"
    taxon_pft_lookup$commonName[i] = "gumbo limbo"}

  else if(genusSpecies == "Carpinus caroliniana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "American hornbeam"}

  else if(genusSpecies == "Carya aquatica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "water hickory"}

  else if(genusSpecies == "Carya cordiformis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "bitternut hickory"}

  else if(genusSpecies == "Carya glabra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "pignut hickory"}

  else if(genusSpecies == "Carya illinoinensis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "pecan"}

  else if(genusSpecies == "Carya ovalis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "red hickory"}

  else if(genusSpecies == "Carya ovata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "shagbark hickory"}

  else if(genusSpecies == "Carya tomentosa"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "mockernut hickory"}

  else if(genusSpecies == "Celtis laevigata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sugarberry"}

  else if(genusSpecies == "Celtis occidentalis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "common hackberry"}

  else if(genusSpecies == "Cercis canadensis"){
    taxon_pft_lookup$pft[i] = "???"   # ???
    taxon_pft_lookup$commonName[i] = "eastern redbud"}

  else if(genusSpecies == "Chrysothamnus viscidiflorus"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "yellow rabbitbrush"}

  else if(genusSpecies == "Cordia collococca"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "red manjack"}

  else if(genusSpecies == "Cornus drummondii"){
    taxon_pft_lookup$pft[i] = "???"    # ???
    taxon_pft_lookup$commonName[i] = "roughleaf dogwood"}

  else if(genusSpecies == "Cornus florida"){
    taxon_pft_lookup$pft[i] = "???"    # ???
    taxon_pft_lookup$commonName[i] = "flowering dogwood"}

  else if(genusSpecies == "Corylus cornuta"){
    taxon_pft_lookup$pft[i] = "woody_shrub"      # deciduous shrubby hazel
    taxon_pft_lookup$commonName[i] = "beaked hazelnut"}

  else if(genusSpecies == "Diospyros virginiana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "common persimmon"}

  else if(genusSpecies == "Ephedra trifurca"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "longleaf jointfir"}

  else if(genusSpecies == "Ericameria laricifolia"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "turpentine bush"}

  else if(genusSpecies == "Exostema caribaeum"){
    taxon_pft_lookup$pft[i] = "woody_shrub"  # "evergreen shrub or small tree" 
    taxon_pft_lookup$commonName[i] = "Caribbean princewood"}

  else if(genusSpecies == "Fagus grandifolia"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "American beech"}

  else if(genusSpecies == "Frangula purshiana"){
    taxon_pft_lookup$pft[i] = "???"     # ???
    taxon_pft_lookup$commonName[i] = "Cascara buckthorn"}

  else if(genusSpecies == "Fraxinus americana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "white ash"}

  else if(genusSpecies == "Fraxinus nigra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "black ash"}

  else if(genusSpecies == "Fraxinus pennsylvanica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "green ash"}

  else if(genusSpecies == "Fraxinus sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "ash"}

  else if(genusSpecies == "Gaultheria shallon"){
    taxon_pft_lookup$pft[i] = "woody_shrub"     # broadleaved evergreen shrub
    taxon_pft_lookup$commonName[i] = "salal"}

  else if(genusSpecies == "Gleditsia triacanthos"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "honey locust"}

  else if(genusSpecies == "Gutierrezia sarothrae"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "broom snakeweed"}

  else if(genusSpecies == "Halesia sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"  # small, deciduous, understory tree 
    taxon_pft_lookup$commonName[i] = "silverbell"}

  else if(genusSpecies == "Hamamelis virginiana"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "American witchhazel"}

  else if(genusSpecies == "Ilex montana"){
    taxon_pft_lookup$pft[i] = "???"          # ???
    taxon_pft_lookup$commonName[i] = "mountain holly"}

  else if(genusSpecies == "Ilex opaca"){
    taxon_pft_lookup$pft[i] = "woody_shrub"   # ???      # broadleaved evergreen tree
    taxon_pft_lookup$commonName[i] = "American holly"}

  else if(genusSpecies == "Juglans nigra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "black walnut"}

  else if(genusSpecies == "Juniperus osteosperma"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Utah juniper"}

  else if(genusSpecies == "Juniperus virginiana"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "red cedar"}

  else if(genusSpecies == "Larix sp."){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "larch"}

  else if(genusSpecies == "Larrea tridentata"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "creosote bush"}

  else if(genusSpecies == "Liquidambar styraciflua"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sweetgum"}

  else if(genusSpecies == "Liriodendron tulipifera"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "tuliptree"}

  else if(genusSpecies == "Lonicera maackii"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "Amur honeysuckle"}

  else if(genusSpecies == "Maclura pomifera"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"   # small deciduous tree or large shrub
    taxon_pft_lookup$commonName[i] = "Osage-orange"}

 else if(genusSpecies == "Magnolia sp."){
    taxon_pft_lookup$pft[i] = "???"
    taxon_pft_lookup$commonName[i] = "magnolia"}

  else if(genusSpecies == "Mahonia nervosa"){
    taxon_pft_lookup$pft[i] = "woody_shrub"      # evergreen shrub
    taxon_pft_lookup$commonName[i] = "Cascade barberry"}

  else if(genusSpecies == "Mangifera indica"){
    taxon_pft_lookup$pft[i] = "???"       # ???   # broadleaf evergreen tree
    taxon_pft_lookup$commonName[i] = "mango"}

  else if(genusSpecies == "Melia azedarach"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "Chinaberrytree"}

  else if(genusSpecies == "Morus rubra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "red mulberry"}

  else if(genusSpecies == "Nyssa aquatica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "water tupelo"}

  else if(genusSpecies == "Nyssa biflora"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "swamp tupelo"}

  else if(genusSpecies == "Nyssa sylvatica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "blackgum"}

  else if(genusSpecies == "Oxydendrum sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sourwood"}

  else if(genusSpecies == "Picea engelmannii"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Engelmann spruce"}

  else if(genusSpecies == "Picea glauca"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "white spruce"}

  else if(genusSpecies == "Picea mariana"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "black spruce"}

  else if(genusSpecies == "Picea sp."){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "spruce"}

  else if(genusSpecies == "Pinus contorta"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "lodgepole pine"}

  else if(genusSpecies == "Pinus echinata"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "shortleaf pine"}

  else if(genusSpecies == "Pinus flexilis"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "limber pine"}

  else if(genusSpecies == "Pinus palustris"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "longleaf pine"}

  else if(genusSpecies == "Pinus ponderosa"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "ponderosa pine"}

  else if(genusSpecies == "Pinus sp."){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "pine"}

  else if(genusSpecies == "Pinus strobus"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "eastern white pine"}

  else if(genusSpecies == "Pinus taeda"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "loblolly pine"}

  else if(genusSpecies == "Pisonia albida"){
    taxon_pft_lookup$pft[i] = "???"    # Growth Form - Multiple Stem, Height, Mature - 30ft
    taxon_pft_lookup$commonName[i] = "corcho bobo"}

  else if(genusSpecies == "Platanus occidentalis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "American sycamore"}

  else if(genusSpecies == "Populus deltoides"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "eastern cottonwood"}

  else if(genusSpecies == "Populus grandidentata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "bigtooth aspen"}

  else if(genusSpecies == "Populus sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "cottonwood"}

  else if(genusSpecies == "Populus tremuloides"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "quaking aspen"}

  else if(genusSpecies == "Prosopis glandulosa"){
    taxon_pft_lookup$pft[i] = "???"   # shrub or tree ranging from 4 to 6m tall
    taxon_pft_lookup$commonName[i] = "honey mesquite"}

  else if(genusSpecies == "Prosopis velutina"){
    taxon_pft_lookup$pft[i] = "???" # deciduous shrub or tree is less than 15 m
    taxon_pft_lookup$commonName[i] = "velvet mesquite"}

  else if(genusSpecies == "Prunus americana"){
    taxon_pft_lookup$pft[i] = "???" # deciduous large shrub or small tree, reaching heights up to 15 feet
    taxon_pft_lookup$commonName[i] = "American plum"}

  else if(genusSpecies == "Prunus angustifolia"){
    taxon_pft_lookup$pft[i] = "woody_shrub" # short, thickly branched shrub 3 to 10ft tall
    taxon_pft_lookup$commonName[i] = "Chickasaw plum"}

  else if(genusSpecies == "Prunus mexicana"){
    taxon_pft_lookup$pft[i] = "???"
    taxon_pft_lookup$commonName[i] = "Mexican plum"}

  else if(genusSpecies == "Prunus serotina"){
    taxon_pft_lookup$pft[i] = "???"       
    taxon_pft_lookup$commonName[i] = "black cherry"}

  else if(genusSpecies == "Pseudotsuga menziesii"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Douglas-fir"}

  else if(genusSpecies == "Quercus alba"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "white oak"}

  else if(genusSpecies == "Quercus coccinea"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "scarlet oak"}

  else if(genusSpecies == "Quercus falcata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "southern red oak"}

  else if(genusSpecies == "Quercus geminata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sand live oak"}

  else if(genusSpecies == "Quercus hemisphaerica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "Darlington oak"}

  else if(genusSpecies == "Quercus laevis"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "turkey oak"}

  else if(genusSpecies == "Quercus laurifolia"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "laurel oak"}

  else if(genusSpecies == "Quercus lyrata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "overcup oak"}

  else if(genusSpecies == "Quercus macrocarpa"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "bur oak"}

  else if(genusSpecies == "Quercus margaretta"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sand post oak"}

  else if(genusSpecies == "Quercus marilandica"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "blackjack oak"}

  else if(genusSpecies == "Quercus michauxii"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "swamp chestnut oak"}

  else if(genusSpecies == "Quercus montana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "chestnut oak"}

  else if(genusSpecies == "Quercus muehlenbergii"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "chinquapin oak"}

  else if(genusSpecies == "Quercus nigra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "water oak"}

  else if(genusSpecies == "Quercus pagoda"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "cherrybark oak"}

  else if(genusSpecies == "Quercus phellos"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "willow oak"}

  else if(genusSpecies == "Quercus rubra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "northern red oak"}

  else if(genusSpecies == "Quercus shumardii"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "Shumard's oak"}

  else if(genusSpecies == "Quercus sp."){
    taxon_pft_lookup$pft[i] = "???"  # apparently there are evergreen oak species? 
    taxon_pft_lookup$commonName[i] = "oak"}

  else if(genusSpecies == "Quercus stellata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "post oak"}

  else if(genusSpecies == "Quercus velutina"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "black oak"}

  else if(genusSpecies == "Quercus virginiana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "live oak"}

  else if(genusSpecies == "Rhamnus davurica"){
    taxon_pft_lookup$pft[i] = "???"     #  spreading shrub or small tree reaching 25 to 30 feet
    taxon_pft_lookup$commonName[i] = "Dahurian buckthorn"}

  else if(genusSpecies == "Rhus aromatica"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "fragrant sumac"}

  else if(genusSpecies == "Rhus glabra"){
    taxon_pft_lookup$pft[i] = "woody_shrub"
    taxon_pft_lookup$commonName[i] = "smooth sumac"}

  else if(genusSpecies == "Robinia pseudoacacia"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "black locust"}

  else if(genusSpecies == "Sassafras albidum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest" # deciduous shrub or tree. trees are 9 to 18 m
    taxon_pft_lookup$commonName[i] = "sassafras"}

  else if(genusSpecies ==  "Sassafras sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "sassafras"}

  else if(genusSpecies == "Sideroxylon lanuginosum"){
    taxon_pft_lookup$pft[i] = "???" # deciduous multi-trunked tall shrub or medium-sized tree to 48ft or more
    taxon_pft_lookup$commonName[i] = "gum bully"}

  else if(genusSpecies == "Symplocos sp."){
    taxon_pft_lookup$pft[i] = "???"  # "deciduous or evergreen shrub or tree"
    taxon_pft_lookup$commonName[i] = "sweetleaf"}

  else if(genusSpecies == "Taxus brevifolia"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Pacific yew"}

  else if(genusSpecies == "Tetradium daniellii"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "bee-bee tree"}

  else if(genusSpecies == "Triadica sebifera"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "Chinese tallow"}

  else if(genusSpecies == "Tsuga canadensis"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "eastern hemlock"}

  else if(genusSpecies == "Tsuga heterophylla"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "western hemlock"}

  else if(genusSpecies == "Ulmus alata"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "winged elm"}

  else if(genusSpecies == "Ulmus americana"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "American elm"}

  else if(genusSpecies == "Ulmus crassifolia"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "cedar elm"}

  else if(genusSpecies == "Ulmus rubra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "slippery elm"}

  else if(genusSpecies == "Ulmus sp."){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "elm"}

  else if(genusSpecies == "Vaccinium parvifolium"){ 
    taxon_pft_lookup$pft[i] = "woody_shrub"       # deciduous Shrub
    taxon_pft_lookup$commonName[i] = "red huckleberry"}

  else if(genusSpecies == "Zinnia acerosa"){
    taxon_pft_lookup$pft[i] = "grass_herb" # forb/herb, subshrub - grass_herb?
    taxon_pft_lookup$commonName[i] = "desert zinnia"}

  # For taxons not described in this code
  #else #print("PFT unknown")

}

```

```{r taxon_pft_table}
kableExtra::kable(taxon_pft_lookup) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped")
```

```{r write_csv}
# write the lookup table to a csv 
write_csv(x = taxon_pft_lookup,
          file = here::here("R","output","taxon_pft_lookup.csv"))
```

## R packages: attempting to automate PFT labeling

### USDA Plants Database

USDA Plants Database R API tutorial: https://recology.info/2016/10/usda-plants-database-r/

The USDA maintains a database of plant information, some of it trait data, some of it life history. Check it out at http://plants.usda.gov/java/

```{r install_load_packages, message=FALSE}
#install.packages("request")
#library(devtools)
#devtools::install_github("sckott/request")
library(request)
library(tibble)
library(tidyverse)
library(here)
```

```{r usda_plants}
# Set the base url
root <- api("https://plantsdb.xyz")

root %>% api_path(heartbeat)
```

^^ My results here do not match those that show up on the tutorial page above. Maybe the API is down or the URL has changed?

Based on the GitHub issue discussion (https://github.com/USDA/USDA-APIs/issues/7), someone uploaded USDA plants data here from 2016: https://uofi.app.box.com/v/usdaplants

```{r try_download, eval=FALSE, include=FALSE}
# Specify URL where file is stored 
url <- "https://uofi.app.box.com/v/usdaplants/folder/6688524249/usda_plants_20160713.csv"
destFile <- file.path(here::here(), "analysis","data", "usda_plants_20160713.csv")
# Apply download.file function in R
download.file(url, destFile)
```

```{r, include = FALSE, eval=FALSE}
#usda_plants <- read_csv(file.path(here::here(), "analysis", "data", "usda_plants_20160713.csv"))

# Error in nchar(x) : invalid multibyte string, element 2
# ^ Issue with some of the column names in the .csv? 
```

### BIEN

RBIEN tutorial: https://cmerow.github.io/RDataScience/3_3_RBIEN_tutorial.html

```{r load_packages_bien, message=FALSE, warning=FALSE}
library(BIEN)
library(ape) # Package for working with phylogenies in R
library(maps) # Useful for making quick maps of occurrences
library(sp) # A package for spatial data
```

BIEN traits and taxonomic data

"Let’s say we’re interested in the genus [Abies], and we’d like to get an idea of how many species there are in this genus and what higher taxa it falls within."

*Abies* genus is fir trees (coniferous trees). The "higher_plant_group" returned by the BIEN taxonomy function is "gymnosperms (conifers)"

```{r bien_taxonomy_1}
Abies_taxonomy<-BIEN_taxonomy_genus(genus = "Abies")

#View(Abies_taxonomy)
Abies_taxonomy[1:3,]
```

*Acer rubrum* species is the red maple (deciduous trees). The "higher_plant_group" returned by the BIEN taxonomy is "flowering plants".

```{r bien_taxonomy_2}
Acer_rubrum_taxonomy <- BIEN_taxonomy_species(species = "Acer rubrum")
Acer_rubrum_taxonomy[1:3,]
```

### TRY Database

https://www.try-db.org/TryWeb/Home.php

### Taxize

https://www.rdocumentation.org/packages/taxize/versions/0.9.99

Time the code. How long did this take? 
```{r how_long_did_this_take}
# stop timer
end_time <- Sys.time()

elapsed_time <- end_time - start_time

print(difftime(end_time, start_time, 
         units = c("auto", "secs", "mins", "hours",
                   "days", "weeks")))
```